---
title: "Final Report for IMDb Movie Bias"
author: "Marie Hasegawa - `mhasegawa7045@floridapoly.edu`"
output: html_notebook
---

This is where the final project report write-up goes. 


Before you submit, make sure everything runs as expected.

You can add sections as you see fit. Make sure you have a section called "Introduction" at the beginning and a section called "Conclusion" at the end. The rest is up to you!

## Section 1: Introduction

The box office industry is one of the biggest global markets in the entertainment industry. According to `IMDbPro`,the industry has a yearly average ` box office revenue` of around 11 to 14 million dollars and a yearly `total gross` of around 9 to 12 billion dollars.

The purpose of the project is to see how a movie's`genre`, `runtime`, `director`, and `actors` can affect the success of a movie and create bias ratings, which will be based on the given `rating`, `votes`, `revenue`, and `metascore`. This project will be using the `IMDB_movies.csv` dataset, which is available in this link, [IMDB_movies.csv](https://github.com/reisanar/datasets/blob/master/IMDB_movies.csv) and in `Section 2: Data Analysis`.


## Section 2: Data Analysis

The dataset, `IMDB_movies.csv`, is a collection of popular movies rated on IMDb for the past decade. The data fields included are `Rank`, `Title`, `Genre`, `Description`, `Director`, `Actors`, `Year`, `Runtime (Minutes)`, `Rating`, `Votes`, `Revenue (Millions)`, and `Metascore`. `IMDB_movies.csv` was provided by Professor Rei Sanchez-Arias under his GitHub account, `reisanar`. The dataset only contains 1000 movies (rows) from 2006 to 2016, so movies past 2006 and from the 20th century will take no part in this analysis. 




- Finding out how `ratings`, `votes`, and `metascore` are affected by who the `director` and `actors` are in the movie by analyzing each `director's` and `actor's` frequency of highly rated movies.

Ratings can be affected by who directed the movie and who starred in it. Directors and actors can create bias in `ratings` that conflict with `metascores`. This hypothesis will be tested by analyzing each `director's` and `actor's` frequency of highly rated movies.

Based on average ratings and highest frequency of  

```{r}
num_movies_by_director<-tally(group_by(IMDB,Director), sort=TRUE) %>%
  head(n=10)
ggplot(data = num_movies_by_director, mapping = aes(x = reorder(Director, n), n)) +
  ggtitle("Director's Movie Count w/ Highest Ratings") +
  geom_bar(stat = "identity", color="black", fill="green") + coord_flip()

IMDB_with_tally <- IMDB %>%
      arrange(desc(Rating)) %>%
      select(Title, Director, Actors, Rating, Metascore) %>% 
      group_by(Director) %>% 
      mutate(count_mov = n())
```
```{r}
avg_rating_table<- IMDB %>% 
  group_by(Director) %>%
  summarize(avg_rating=mean(Rating, na.rm=TRUE)) %>%
  arrange(desc(avg_rating)) %>%
  head(n=10)

ggplot(data = avg_rating_table, mapping = aes(x = reorder(Director, avg_rating), avg_rating)) + 
  geom_bar(stat = "identity", color="black", fill="pink") + coord_flip()
```
Out of all directors, `Christopher Nolan` is the most frequent director with the highest ratings from his movies: `The Dark Knight`, `Inception	`, `Interstellar`, `The Prestige`, and `The Dark Knight Rises`.  


- Deduce if `rating` and `metascore` correlates with each other.
```{r}
RateVsMeta<-IMDB %>%
  select(Title, Rating, Metascore) %>%
  arrange(desc(Rating))
RateVsMeta

MetaVsRate<- IMDB %>%
  select(Title, Rating, Metascore) %>%
  arrange(desc(Metascore))
MetaVsRate
```
- See if a high `revenue` will lead to a high `rating`, `votes`, and `metascore`, and vice versa.





















